\bfseries\LARGE\singlespacing{Brief report: SARS-CoV-2 more frequently detected in indoor air than other respiratory viruses}

### 2023-12-04 doc/comparison-mcid12/manuscript.tex

\begin{information}\normalfont
\textbf{Word count:} Abstract 252, Manuscript 3,244 \\
\textbf{Display items:} Figure 1 \\
\textbf{References:} 37 \\
\textbf{Supporting information:} STROBE checklist, Supplementary Data with texts, tables, and figures
% \par
\end{information}

\begin{abstract}\normalfont
\noindent We compared air, air filters, and saliva samples from two studies conducted in winter 2021/2022 (during the SARS-CoV-2 omicron wave) and winter 2022/2023 (after the COVID-19 pandemic) in a Swiss school setting. Whereas during the pandemic we detected almost exclusively SARS-CoV-2 in human saliva (19~samples of SARS-CoV-2, 2~samples of non-SARS-CoV-2), the most common respiratory viruses after the pandemic were adenovirus, influenza~B, and rhinoviruses (3~samples of SARS-CoV-2, 47~samples of non-SARS-CoV-2). Despite the same study setting and environmental conditions, airborne detection of SARS-CoV-2 was relatively more frequent than the detection of non-SARS-CoV-2 viruses (10~samples of SARS-CoV-2 in 2021/2022 vs 2~samples of non-SARS-CoV-2 in 2022/2023). Our comparison suggests that respiratory viruses other than SARS-CoV-2 may be less detectable in the air and that close contact is required for transmission. \medskip

@@

%\section*{Introduction}

@@

\bibliography{../paper/references.bib}

>>>>

\begin{information}\normalfont
\textbf{Word count:} Abstract 126 (max 150), Manuscript xx (max 1,500) \\
\textbf{Display items:} Figure 1 \\
\textbf{References:} xx (max 12) \\
%\textbf{Supporting information:} STROBE checklist, Supplementary Data with texts, tables, and figures
% \par
\end{information}

\begin{abstract}\normalfont
\noindent We compare air, air filters, and saliva samples from two studies conducted in winter 2021/2022 (during the SARS-CoV-2 omicron wave) and winter 2022/2023 (after the COVID-19 pandemic) in a Swiss school setting. Whereas during the pandemic we detected almost exclusively SARS-CoV-2 in human saliva (19~samples of SARS-CoV-2, 2~samples of non-SARS-CoV-2), the most common respiratory viruses after the pandemic were adenovirus, influenza~B, and rhinoviruses (3~samples of SARS-CoV-2, 47~samples of non-SARS-CoV-2). Despite the same study setting and environmental conditions, airborne detection of SARS-CoV-2 was relatively more frequent than the detection of non-SARS-CoV-2 viruses (10~samples of SARS-CoV-2 in 2021/2022 vs 2~samples of non-SARS-CoV-2 in 2022/2023). Our comparison suggests that respiratory viruses other than SARS-CoV-2 may be less detectable in the air and that close contact is required for transmission. \medskip

@@

Respiratory viruses, such as COVID-19 and influenza, pose significant challenges in terms of transmission within schools and and other indoor settings\cite{Leung2020}. These viruses spread through various routes, including respiratory particles like large droplets and small aerosols. Unlike larger droplets that settle quickly, aerosols can remain suspended in the air for extended periods\cite{Wang2020,Morawska2020}, contributing to the overall transmission dynamics. Detecting respiratory viruses in the air is challenging\cite{Wang2021,Belser2023PLOSPath}, adding to the complexity of understanding their spread.

Infectious pathogens are primarily found in smaller particles (<5$\mu$m) and the distribution is similar across a range of different pathogens\cite{Fennelly2020}. Thus, infectious aerosols have the potential for long-range transmission, but the larger concentration of particles near the infectious person favours short-range transmission\cite{Tellier2009JTRSI,Wang2020,Morawska2020}. However, previous work also suggests that prolonged close contact may facilitate transmission\cite{Leung2020NatMed,Brankston2007LancetID,Lind2023NatCommun}. The complexity of respiratory virus transmission lies in the interplay between the physicochemical properties of the virus-laden aerosols and environmental conditions\cite{Wang2020,Morawska2020}, making it challenging to devise effective public health strategies for mitigating the spread of respiratory infections in various indoor settings, including schools.

In this study, we compare samples from human saliva, the air, and the filters of air cleaners from two studies conducted in winter 2021/2022 (during the SARS-CoV-2 omicron wave)\cite{Banholzer2023PLoSMed} and winter 2022/2023 (after the COVID-19 pandemic)\cite{Banholzer2023submitted} in a Swiss school setting. We show differences in the distribution of respiratory viruses and emphasise the varying ability to detect these viruses in the air.

@@

\bibliography{references.bib}
